Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-03-26 introduction.tex
introduction. first quick pass.

>>>>

As the LHC prepares for the high-luminosity era (Run 3 in $~\approx$ 2022), it
is anticipated that the data volumes that will need analyzing will increase by
10X compared to the current phase (Run 2).   The data will be larger in volume
but will also require heterogeneous computational processes. In spite of the
impressive scale of the ATLAS distributed computing system, demand for
computing systems will significantly outstrip supply (availability).

There are multiple levels at which this problem needs to be addressed
urgently, e.g., the utilization of emerging parallel architectures (e.g.,
platforms), algorithmic and advances in analytical methods (e.g., use of
Machine Learning) and the ability to exploit different platforms (e.g., clouds
and supercomputers).

This paper is a case study of how the ATLAS experiment has "broken free" of
the traditional computational approach of high-throughput computing on
distributed resources to embrace new platforms, in particular high-performance
computers. Specifically, we discuss the experience of integrating the PanDA
workload management system with Titan ~\textemdash{} a DOE leadership
computing facility and scaled to analyze up to \jhanote{XX\%} of the ATLAS
experiment workload. We present the design of PanDA and show how localized
substitution to use well established pilot-concepts allow enhanced support for
heterogeneous workloads (such as molecular dynamics) and advanced execution
modes.

This state-of-practice paper provides multiple contributions.  (i) Documents
the many design and operational considerations that have been  taken to
support the sustained, scalable and production usage of Titan for historically
high-throughput workloads, (ii) Extensions to PanDA to support non-traditional
heterogeneous workloads and execution modes,  and (iii) As the community looks
forward to designing the next generation of online analytical platforms
~\cite{foap} the lessons learnt from our project provide some guidance for how
current and future experimental and observational systems can be integrated
with supercomputers in production.
